---
title: "National Cancer Patient Experience Survey (Preparation)"
output: html_notebook
---

In this notebook I load the National Cancer Patient Experience Survey (NCPES) dataset and cluster it using different 
hyperparameters. Note that due to privacy reasons we cannot publicly share the raw data, so you would need to apply
for access yourself [here](https://discover.ukdataservice.ac.uk/Catalogue/?sn=8163&type=Data%20catalogue) to reproduce
the data preparation and clustering results. But due to their aggregated form we can share the results from the clustering 
algorithms (`data/ncpes_clustering_results`), so you can reproduce the subsequent analysis in the `ncpes_analysis.RMD` notebook.

```{r setup}
library(tidyverse)
library(stringr)
library(purrr)
library(mixdir)
library(parallel)
```


```{r}
if(file.exists("../data/ncpes_data/UKDA-8163-tab")){
  print("Data available, reproducing data preparation and clustering")
  warning("It can take days to run the complete notebook. You should probably run it in parallel on a cluster!")
}else{
  stop(paste0("Due to data privacy we cannot share this dataset. Please get if from ",
             "here (https://discover.ukdataservice.ac.uk/Catalogue/?sn=8163&type=Data%20catalogue) to fully reproduce the analysis"))
  
}
```

```{r}
# Load the data
cpes <- read_delim("../data/ncpes_data/UKDA-8163-tab/tab/cpes_2015_eul_file_21-02-2017-datav3.tab", 
                   "\t", escape_double = FALSE, trim_ws = TRUE)

# Remove all Score columns and other questions that are problematic
X <- cpes %>%
  dplyr::select(- contains("Score"), -DateReceived, -Sample_LengthOfStay, - matches("^Q5[a-z]$"),
         - matches("^Q25[a-z]$"), -Q59, -Q64, -IMD_Quintile, - ICD10BasedCancerType, -Ethnicity) %>%
  as.matrix() %>%
  apply(2, as.character)

```


# Dir 

```{r}
ks <- rep(c(5,50, 100, 200), each=5)


cluster_list <- mclapply(ks, mc.cores=5, function(k){
  mixdir(X, n_latent=k, alpha=c(1,1), select_latent=FALSE, verbose=TRUE,
         max_iter=5000, na_handle="ignore", repetitions=1)
})


write_rds(cluster_list, "../data/ncpes_clustering_results/dir_ks_list.RDS")
```


```{r}
# Calculate the subsampling
ks <- rep(c(5, 10, 15, 20, 25, 30, 40), each=5)


cluster_list <- parallel::mclapply(ks, mc.cores=5, function(k){
  mixdir(X, n_latent=k, alpha=c(1,1), select_latent=FALSE, verbose=TRUE,
         max_iter=5000, na_handle="ignore", repetitions=1)
})


write_rds(cluster_list, "../data/ncpes_clustering_results/dir_ks_list2.RDS")
```


# DP

```{r}
ks <- rep(c(5,50, 100, 200), each=5)


cluster_list <- parallel::mclapply(ks, mc.cores=5, function(k){
  mixdir(X, n_latent=k, alpha=c(1,1), select_latent=TRUE, verbose=TRUE,
         max_iter=5000, na_handle="ignore", repetitions=1)
})


write_rds(cluster_list, "../data/ncpes_clustering_results/dp_ks_list.RDS")
```


```{r}
ks <- rep(c(5, 10, 15, 20, 25, 30, 40), each=5)


cluster_list <- parallel::mclapply(ks, mc.cores=5, function(k){
  mixdir(X, n_latent=k, alpha=c(1,1), select_latent=TRUE, verbose=TRUE,
         max_iter=5000, na_handle="ignore", repetitions=1)
})


write_rds(cluster_list, "../data/ncpes_clustering_results/dp_ks_list2.RDS")

```


# Alphas

```{r}
alphas <-rep(c(1, 10, 100, 500, 1000, 1e4, 1e5), each=5)

cluster_list <- parallel::mclapply(alphas, mc.cores=15, function(a){
  mixdir(X, n_latent=15, alpha=c(a,1), select_latent=TRUE, verbose=TRUE,
         max_iter=400, na_handle="ignore", repetitions=1)
})


write_rds(cluster_list, "../data/ncpes_clustering_results/dp_alphas_list3.RDS")
```












































