---
title: "Mental Health in Tech Survey 2016"
output: html_notebook
---


In this notebook I want to analyze the mental health in tech 2016 survey.



```{r setup}
library(tidyverse)
library(mixdir)
library(cowplot)
```



Load data

Before we can actually use the data we need to clean up the data a little bit:

- Self employed people are asked different questions (-> lots of missing values)
- Disorder is a list with entries, which are separated by a |
- Gender question is a free form text field, with some _unusual_ answers
- Other questions are free form
- position is also multiple choice
- If living in US, US states are specifically asked for
- Age sometimes non-sense


```{r}
raw_data <- read_csv("../data/osmi_mental_health/mental-heath-in-tech-2016_20161114.csv")
cleaned_data  <- raw_data %>%  
  # Remove free form columns:
  dplyr::select(- `Why or why not?`, - `Why or why not?_1`) %>%
  # Remove outliers in age column and create age groups
  mutate(`What is your age?` = if_else(`What is your age?` > 70 | `What is your age?` < 20, NA_integer_, `What is your age?`)) %>%
  mutate(`What is your age?` = cut(`What is your age?`, breaks=5)) %>%
  # Handle Gender
  mutate(`What is your gender?` = case_when(
    glue::trim(tolower(`What is your gender?`)) %in% c('male', 'm', 'man', 'cis male',
                                                       'male.', 'male (cis)', 'man', 'sex is male',
                                                       'cis male', 'malr', 'dude', 
                                                       "i'm a man why didn't you make this a drop down question. you should of asked sex? and i would of answered yes please. seriously how much text can this take?",
                                                       'mail', 'm|', 'male (trans, ftm)',
                                                       'cisdude', 'cis man') ~ "male",
    glue::trim(tolower(`What is your gender?`)) %in% c('female', 'i identify as female.', 'female',
                                                       'female assigned at birth', 'f', 'woman', 'fm', 'f',
                                                       'cis female', 'transitioned, m2f', 'female or multi-gender femme',
                                                       'female/woman', 'cisgender female', 
                                                       'mtf', 'fem', 'female (props for making this a freeform field, though)',
                                                       'cis-woman', 'afab', 'transgender woman') ~ "female",
    glue::trim(tolower(`What is your gender?`)) %in% c("androgynous", "bigender", "enby", "other/transfeminine", "agender",
                                                       "female-bodied; no feelings about gender" , "fluid", "genderfluid", 
                                                       "genderfluid (born female)", "genderflux demi-girl", "genderqueer",
                                                       "genderqueer woman", "human", "male 9:1 female, roughly", "male/genderqueer",
                                                       "nb masculine", "non-binary", "nonbinary", "queer") ~ "queer",
    glue::trim(tolower(`What is your gender?`)) %in% c("n/a", "none of your business", "other", "unicorn") ~ NA_character_,
    TRUE ~glue::trim(tolower(`What is your gender?`))
  )) %>%
  # Remove US State question and work
  dplyr::select(- `What US state or territory do you live in?`, - `What US state or territory do you work in?`, - `What country do you work in?`) %>%
  # Simplifying the country question
  mutate(`What country do you live in?`= forcats::fct_lump(`What country do you live in?`, n=3)) %>%
  # Split position
  mutate(`Which of the following best describes your work position?`= stringr::str_split(`Which of the following best describes your work position?`, "\\|")) %>%
  rowwise() %>%
  mutate(`Which of the following best describes your work position?`=case_when(
    "Executive Leadership" %in% `Which of the following best describes your work position?` ~ "leadership",
    "Supervisor/Team Lead" %in% `Which of the following best describes your work position?` ~ "leadership",
    "One-person shop" %in% `Which of the following best describes your work position?` ~ "one-person shop",
    "Back-end Developer"  %in% `Which of the following best describes your work position?` ~ "back-end developer",
    "Front-end Developer"  %in% `Which of the following best describes your work position?` ~ "front-end developer",
    TRUE ~ "other"
  )) %>%
  # Fixing the disorders
  mutate(`If yes, what condition(s) have you been diagnosed with?`= stringr::str_split(`If yes, what condition(s) have you been diagnosed with?`, "\\|")) %>%
  mutate(`If yes, what condition(s) have you been diagnosed with?`=case_when(
    any(c("Autism (Asperger's)", "Autism", "Asperges", "Autism Spectrum Disorder", "PDD-NOS") %in% `If yes, what condition(s) have you been diagnosed with?`) ~ "autism",
    any(c("Combination of physical impairment (strongly near-sighted) with a possibly mental one (MCD / \"ADHD\", though its actually a stimulus filtering impairment)",
          "ADD (w/o Hyperactivity)", "Attention Deficit Hyperactivity Disorder") %in% `If yes, what condition(s) have you been diagnosed with?`) ~ "adhd",
    any(c("Post-traumatic Stress Disorder", "Stress Response Syndromes") %in% `If yes, what condition(s) have you been diagnosed with?`) ~ "ptsd",
    "Mood Disorder (Depression, Bipolar Disorder, etc)" %in% `If yes, what condition(s) have you been diagnosed with?` ~ "mood disorder",
    "Depression" %in% `If yes, what condition(s) have you been diagnosed with?` ~ "mood disorder",
    "Anxiety Disorder (Generalized, Social, Phobia, etc)" %in% `If yes, what condition(s) have you been diagnosed with?` ~ "anxiety disorder",
    "I haven't been formally diagnosed, so I felt uncomfortable answering, but Social Anxiety and Depression." %in% `If yes, what condition(s) have you been diagnosed with?` ~ "anxiety disorder",
    "Substance Use Disorder" %in% `If yes, what condition(s) have you been diagnosed with?` ~ "substance use disorder",
    TRUE ~ paste0(`If yes, what condition(s) have you been diagnosed with?`, collapse = "|")
  )) %>%
  mutate(`If maybe, what condition(s) do you believe you have?`= stringr::str_split(`If maybe, what condition(s) do you believe you have?`, "\\|")) %>%
  mutate(`If maybe, what condition(s) do you believe you have?`=case_when(
    any(c("Asperger Syndrome", "Asperger Syndrome", "Autism") %in% `If maybe, what condition(s) do you believe you have?`) ~ "autism",
    any(c("Attention Deficit Hyperactivity Disorder") %in% `If maybe, what condition(s) do you believe you have?`) ~ "adhd",
    any(c("Post-traumatic Stress Disorder", "Stress Response Syndromes") %in%`If maybe, what condition(s) do you believe you have?`) ~ "ptsd",
    "Mood Disorder (Depression, Bipolar Disorder, etc)" %in% `If maybe, what condition(s) do you believe you have?` ~ "mood disorder",
    "Anxiety Disorder (Generalized, Social, Phobia, etc)" %in% `If maybe, what condition(s) do you believe you have?` ~ "anxiety disorder",
    any(c("Substance Use Disorder", "Addictive Disorder") %in% `If maybe, what condition(s) do you believe you have?`) ~ "substance use disorder",
    TRUE ~ paste0(`If maybe, what condition(s) do you believe you have?`, collapse = "|")
  )) %>%
  mutate(`If so, what condition(s) were you diagnosed with?`= stringr::str_split(`If so, what condition(s) were you diagnosed with?`, "\\|")) %>%
  mutate(`If so, what condition(s) were you diagnosed with?`=case_when(
    any(c("Asperger Syndrome", "Asperger Syndrome", "Autism", "Autism (Asperger's)",
          "PDD-NOS (see above)", "PDD-NOS", "Autism - while not a \"mental illness\", still greatly affects how I handle anxiety",
          "Aspergers", "autism spectrum disorder") %in% `If so, what condition(s) were you diagnosed with?`) ~ "autism",
    any(c("Attention Deficit Hyperactivity Disorder", "MCD (when it was diagnosed, the ultra-mega \"disorder\" ADHD didn't exist yet)",
          "attention deficit disorder (but not the hyperactive version)", "Autism Spectrum Disorder", "ADD (w/o Hyperactivity)") %in% `If so, what condition(s) were you diagnosed with?`) ~ "adhd",
    any(c("Post-traumatic Stress Disorder", "Stress Response Syndromes") %in%`If so, what condition(s) were you diagnosed with?`) ~ "ptsd",
    "Mood Disorder (Depression, Bipolar Disorder, etc)" %in% `If so, what condition(s) were you diagnosed with?` ~ "mood disorder",
    "Depression" %in% `If yes, what condition(s) have you been diagnosed with?` ~ "mood disorder",
    "Anxiety Disorder (Generalized, Social, Phobia, etc)" %in% `If so, what condition(s) were you diagnosed with?` ~ "anxiety disorder",
    any(c("Substance Use Disorder", "Addictive Disorder") %in% `If so, what condition(s) were you diagnosed with?`) ~ "substance use disorder",
    TRUE ~ paste0(`If so, what condition(s) were you diagnosed with?`, collapse = "|")
  )) %>%
  ungroup() %>%
  # Join the responses which mental disorder they have from yes and maybe people
  mutate(mental_health_condition=case_when(
    ! is.na(`If yes, what condition(s) have you been diagnosed with?`) ~ `If yes, what condition(s) have you been diagnosed with?`,
    ! is.na(`If maybe, what condition(s) do you believe you have?`) ~ `If maybe, what condition(s) do you believe you have?`,
    TRUE ~ "not applicable"
  )) %>%
  dplyr::select(- c(`If yes, what condition(s) have you been diagnosed with?`, `If maybe, what condition(s) do you believe you have?`)) %>%
  # Remove irrelevant questions
  dplyr::select(- c(`Is your primary role within your company related to tech/IT?`, `Is your employer primarily a tech company/organization?`))
  

```


```{r}
employed_data <- cleaned_data %>%
  filter(`Are you self-employed?` == 0) %>%
  select_if(~ length(unique(.x)) > 1)
```



# Cluster data

This chunk can take a long time to run. It is a good idea to run it in parallel on a cluster.

```{r}
set.seed(1)
alphas <- c(1, 5, 10, 20, 50, 100, 250, 500, 1000)
cluster_list <- parallel::mclapply(alphas, mc.cores=10, function(a){
  mixdir(X = employed_data, n_latent = 25, select_latent =TRUE, alpha = c(a, 1),
                 max_iter=200, na_handle = "ignore", verbose=TRUE, repetitions = 5)
})
write_rds(cluster_list, "../data/osmi_mental_health/cluster_list.RDS")
```


```{r}
cluster_list <- readRDS("../data/osmi_mental_health/cluster_list.RDS")
res_dp <- cluster_list[alphas == 100][[1]]
```



# Check significance

```{r}
employed_data %>%
  mutate(health_care_coverage=`Do you know the options for mental health care available under your employer-provided coverage?`) %>%
  mutate(negative_consequence=factor(.$`Do you think that discussing a mental health disorder with your employer would have negative consequences?`,
                                    levels=c("Yes", "Maybe", "No"), ordered = TRUE)) %>%
  xtabs(~ negative_consequence + health_care_coverage, data=.) %>%
  print() %>%
  chisq.test() 


employed_data %>%
  filter(.$`Do you currently have a mental health disorder?` == "Yes") %>%
  mutate(ptsd = mental_health_condition == "ptsd") %>%
  mutate(NegativeConsequence=factor(.$`Do you think that discussing a mental health disorder with your employer would have negative consequences?`,
                                    levels=c("Yes", "Maybe", "No"), ordered = TRUE)) %>%
  {
    tmp <- table(list(ptsd=.$ptsd, NegativeConsequence=.$NegativeConsequence))
    print(tmp)
    print(tmp / rowSums(tmp), digits = 2)
    fisher.test(tmp)
  } 

# ADHD does not differ
employed_data %>%
  filter(.$`Do you currently have a mental health disorder?` == "Yes") %>%
  mutate(adhd = mental_health_condition == "adhd") %>%
  mutate(NegativeConsequence=factor(.$`Do you think that discussing a mental health disorder with your employer would have negative consequences?`,
                                    levels=c("Yes", "Maybe", "No"), ordered = TRUE)) %>%
  # filter(NegativeConsequence != "Maybe") %>%
  # mutate(NegativeConsequence = forcats::fct_drop(NegativeConsequence)) %>%
  {
    tmp <- table(list(adhd=.$adhd, NegativeConsequence=.$NegativeConsequence))
    print(tmp)
    print(tmp / rowSums(tmp), digits = 2)
    fisher.test(tmp)
  } 

```




# Create Plot (Fig. 11)

```{r fig.height=7, fig.width=15}
# Diverging colorscheme for 3 classes from colorbrewer2.org
color_scheme_ynm <- c("Yes"="#1b9e77", "Maybe"="#d95f02", "No"="#7570b3")

{
  par(xpd = NA, # switch off clipping, necessary to always see axis labels
      bg = "white", # switch off background to avoid obscuring adjacent plots
      oma=c(2, 1, 2, 1) # Set the margins 
  ) 
  purrr::map2_df(cluster_list, alphas, ~ data_frame(class=.x$pred_class, alpha=.y, individual=1:length(.x$pred_class))) %>%
    mutate(alpha = as.factor(alpha),
           class = as.factor(class)) %>%
    right_join(data.frame(diagnosed=employed_data$`Have you been diagnosed with a mental health condition by a medical professional?` == "Yes",
                          individual=1:nrow(employed_data))) %>%
    filter(alpha == 1 | alpha == 10 | alpha == 100 | alpha == 1000) %>%
    group_by(alpha) %>%
    mutate(class = forcats::fct_lump(LETTERS[class], n=5)) %>%
    ungroup() %>%
    mutate(alpha=paste0(alpha)) %>%
    spread(alpha, class) %>%
    group_by_at(vars(matches("\\d+"), diagnosed)) %>%
    summarize(freq=n()) %>%
    {alluvial::alluvial(.[, 1:4], freq=.$freq, 
                        col=if_else(.$diagnosed, color_scheme_ynm["Yes"], color_scheme_ynm["No"]), 
                        border=if_else(.$diagnosed, color_scheme_ynm["Yes"], color_scheme_ynm["No"]), 
                        hide=.$freq < 10, blocks=TRUE, gap.width = 0.2, alpha=0.8,
                        axis_labels=paste0("α₁=", c(1,10,100,1000)))}
   rect(2.8, -0.15, 3.2, 1.05, border="red", lwd=4)
   alluvial_plot <- recordPlot()
}

tmp_heatmap <- res_dp$class_prob[order(res_dp$pred_class), 1:11]
colnames(tmp_heatmap) <- c(LETTERS[1:10], "...")
heatmap_plot <- as.data.frame(tmp_heatmap) %>%
  mutate(Individuals=1:n()) %>%
  gather(Class, value,  -Individuals) %>%
  mutate(Class=factor(Class, levels=c(LETTERS[1:10], "..."), ordered=TRUE)) %>%
  ggplot(aes(x=Class, y=Individuals, fill=value)) + 
    geom_tile() +
    scale_y_reverse(breaks=0,labels="") + 
    viridis::scale_fill_viridis(name="Prob.")


diagnose_plot <- employed_data %>%
  mutate(class=as.factor(LETTERS[res_dp$pred_class])) %>%
  filter(class == "A" | class == "B" | class == "C" | class == "D") %>%
  mutate(diagnosed=`Have you been diagnosed with a mental health condition by a medical professional?`) %>%
  mutate(diagnosed=forcats::fct_lump(as.factor(diagnosed), n=5)) %>%
  group_by(class) %>%
  summarize(res=list(enframe(table(diagnosed)))) %>%
  unnest() %>%
  mutate(name=factor(name, levels=c("Yes", "Maybe", "No"), ordered=TRUE)) %>%
  ggplot(aes(x=class, y=value, fill=name)) +
    geom_bar(position="stack",stat="identity") +
    scale_fill_manual(name="", values=color_scheme_ynm) +
    ggtitle("Mental Disorder Diag.") +
    xlab("Class") + ylab("Count")

neg_exp_plot <- employed_data %>%
  mutate(class=as.factor(LETTERS[res_dp$pred_class])) %>%
  filter(class == "A" | class == "B" | class == "C" | class == "D") %>%
  mutate(negative_consequence=`Do you think that discussing a mental health disorder with your employer would have negative consequences?`) %>%
  mutate(negative_consequence=forcats::fct_lump(as.factor(negative_consequence), n=5)) %>%
  group_by(class) %>%
  summarize(res=list(enframe(table(negative_consequence)))) %>%
  unnest() %>%
  mutate(name=factor(name, levels=c("Yes", "Maybe", "No"), ordered=TRUE)) %>%
  ggplot(aes(x=class, y=value, fill=name)) +
    geom_bar(position="stack",stat="identity") +
    scale_fill_manual(name="", values=color_scheme_ynm) +
    ggtitle("Expect Neg. Cons.")  +
    xlab("Class") + ylab("Count")

leave_plot <- employed_data %>%
  mutate(leaving=factor(.$`If a mental health issue prompted you to request a medical leave from work, asking for that leave would be:`,
                        levels=c("Very easy", "Somewhat easy", "Neither easy nor difficult", "Somewhat difficult", "Very difficult", "I don't know"),
                        ordered = TRUE)) %>%
  mutate(negative_consequence=factor(.$`Do you think that discussing a mental health disorder with your employer would have negative consequences?`,
                                    levels=c("Yes", "Maybe", "No"), ordered = TRUE)) %>%
  xtabs(~ negative_consequence + leaving, data=.) %>%
  as.data.frame() %>%
  group_by(negative_consequence) %>%
  mutate(Freq=Freq/sum(Freq)) %>%
  ungroup() %>%
  mutate(negative_consequence=factor(negative_consequence, levels=c("No", "Maybe", "Yes"), ordered=TRUE)) %>%
  ggplot(aes(x=leaving, y=negative_consequence, fill=Freq)) +
    geom_tile() +
    geom_vline(xintercept = 5.5) + 
    viridis::scale_fill_viridis() + 
    theme(axis.text.x =  element_text(angle= -16, hjust=0.1)) +
    ggtitle("Requesting leave for mental health issue is...")  +
    ylab("Expect Neg.") + xlab("")


left_plots <- plot_grid(alluvial_plot, leave_plot, labels = c('A', 'E'), ncol = 1, rel_heights = c(2, 1))
right_plots <- plot_grid(heatmap_plot, diagnose_plot, neg_exp_plot, labels = c('B', 'C', 'D'), ncol = 1, rel_heights = c(1, 1, 1), align = "v")
plot_grid(left_plots, right_plots, labels = c('', ''), nrow = 1, rel_widths = c(3, 1))

ggsave("../output/mental_health_complete.png", width=15, height=7)

```




















