---
title: "Clustering Performance on Synthetic Data"
output: html_notebook
---

In this notebook I first generate a dataset with a known ground truth and use that to evaluate the performance and runtime
of 5 different clustering algorithm for categorical data:

1. ROCK as implemented in the cba package
2. k-mode as implemented in the klar package
3. EM for latent class model as implemented in the poLCA package
4. mixdir with a Dirichlet prior from our new mixdir package
5. mixdir with a Dirichlet Process prior from our new mixdir package


```{r setup}
library(tidyverse)
library(mixdir)
library(poLCA)
library(cba)
library(cowplot)
```



```{r}

generate_categorical_dataset <- function(n_ind, n_quest, n_cat, n_true_classes, lambda_true=1/n_true_classes,
                                         bb_size=100, bb_suc=0.05, bb_overdis=10){
  
  if(length(n_cat) > 1 && length(n_cat) != n_quest) stop("n_cat must either be length or match the n_quest")
  if(length(n_cat) == 1){
    n_cat <- rep(n_cat, times=n_quest)
  }
  
  # First step generate true class representives
  U_true <- lapply(1:n_quest, function(j)lapply(1:n_true_classes, function(k) rep(NA, n_cat[j]) ))
  for(k in 1:n_true_classes){
    for(j in 1:n_quest){
      U_true[[j]][[k]] <- rmutil::rbetabinom(n_cat[j], size = bb_size, m=bb_suc, s=bb_overdis) + 1
    }
  }
  
  if(length(lambda_true) == 1) {
    lambda_true <- rep(lambda_true, n_true_classes)
  }
  true_latent <- sample(1:n_true_classes, n_ind, replace=TRUE, prob=lambda_true) 
  
  
  data <- matrix(NA, ncol=n_quest, nrow=n_ind)
  for(i in 1:n_ind){
    for(j in 1:n_quest){
      data[i, j] <- which(extraDistr::rdirmnom(1, 1, alpha=U_true[[j]][[true_latent[i]]]) == 1)
    }
  }
  
  list(X=data, true_latent=true_latent, U_true=U_true)

}

test_cluster_perf <- function(n_iter, n_latent){
  set.seed(1)
  bind_rows(lapply(1:n_iter, function(idx){
    
    tmp <- generate_categorical_dataset(n_ind=1000, n_quest=5, n_cat=rbinom(1, 30, 0.3), n_true_classes = 4, 
                                        lambda_true=extraDistr::rdirichlet(1, rep(1,4)),
                                        bb_size=100, bb_suc=rbeta(1, 0.5, 10), bb_overdis=max(0.1, rnorm(1, 10, 5)))
    X <- tmp$X
    true_classes <- tmp$true_latent
    
    t_md <- system.time(mixdir_result <- mixdir(X, n_latent=n_latent, select_latent = FALSE, alpha=1, beta = 0.1))
    t_dp <- system.time(mixdirdp_result <- mixdir(X, n_latent=n_latent, select_latent = TRUE, alpha=c(1,1), beta = 0.1))
    t_km <- system.time(kmode_result <- klaR::kmodes(X,n_latent))
    t_ro <- system.time(rock_result <- fitted(rockCluster(as.dummy(X), n=n_latent, theta = 0.5, debug=FALSE)))
    t_po <- system.time(lca_result <- poLCA(formula(paste0("cbind(", paste0(paste0("V", 1:ncol(X)), collapse = ","), ") ~ 1")),
                                           as.data.frame(X), nclass=n_latent, verbose = FALSE))
    
    annotation <- data.frame(Group= as.factor(true_classes))
    clustering_matrix <-rbind(kmode_result$cluster,
                              mixdir_result$pred_class,
                              mixdirdp_result$pred_class,
                              lca_result$predclass, as.numeric(rock_result$cl),
                              rep(1, length(true_classes)),
                              true_classes)
    
    arand_idx <- clustering_matrix %>%
      apply(1, function(row) mcclust::arandi(true_classes, row))
    
    tribble(   ~method,       ~time,
             "k_mode",    t_km["user.self"],
             "mixdir",  t_md["user.self"],
             "middir_dp",  t_dp["user.self"],
             "poLCA",      t_po["user.self"],
             "rock",       t_ro["user.self"],
             "all_one",    0,
             "correct",    0) %>%
      mutate(arand_idx=arand_idx,
             iteration=idx)
    
  }))
}

```

```{r}
set.seed(1)
data <- test_cluster_perf(n_latent=4, n_iter=100)
data2 <- test_cluster_perf(n_latent=8, n_iter=100)
```


```{r}
write_rds(list(data=data, data2=data2), "../data/synthetic_clustering_results/data_k_4_vs_8_niter100.RDS")
```

```{r}
data <- read_rds("../data/synthetic_clustering_results/data_k_4_vs_8_niter100.RDS")$data
data2 <- read_rds("../data/synthetic_clustering_results/data_k_4_vs_8_niter100.RDS")$data2
```



```{r}


p1 <- data %>%
  dplyr::select(-time) %>%
  filter(method != "correct" & method != "all_one") %>%
  mutate(method= forcats::fct_reorder(as.factor(method), - arand_idx, mean)) %>%
  ggplot(aes(x=method, y=arand_idx)) +
    ggbeeswarm::geom_quasirandom(width = 0.2) +
    stat_summary(fun.data = "mean_cl_boot", colour = "red", geom = "crossbar", size=0.2, width=0.5) +
    ggsignif::geom_signif(comparisons = list(c("middir_dp", "mixdir"), 
                                             c("middir_dp", "poLCA"), c("k_mode", "rock"), c("mixdir", "k_mode")),
                          y_position =  c(1.16, 1.1, 0.8, 1.0), test.args=list(paired=TRUE), map_signif_level = TRUE, 
                          textsize = 4) +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) +
    scale_y_continuous("ARI", breaks=c(0,0.25, 0.5, 0.75, 1), limits = c(NA, 1.3)) +
    scale_x_discrete("", labels=c(k_mode="k-mode", middir_dp="mixdir DP", mixdir="mixdir", poLCA="poLCA", rock="ROCK")) +
    ggtitle("True Classes = 4 and K=4")

p2 <- data2 %>%
  dplyr::select(-time) %>%
  filter(method != "correct" & method != "all_one") %>%
  mutate(method= forcats::fct_reorder(as.factor(method), - arand_idx, mean)) %>%
  ggplot(aes(x=method, y=arand_idx)) +
    ggbeeswarm::geom_quasirandom(width = 0.2) +
    stat_summary(fun.data = "mean_cl_boot", colour = "red", geom = "crossbar", size=0.2, width=0.5) +
    ggsignif::geom_signif(comparisons = list(c("poLCA", "k_mode"), c("middir_dp", "mixdir"), 
                                             c("mixdir", "poLCA"), c("k_mode", "rock")),
                          y_position = c(0.9, 0.9, 1.0, 0.8), test.args=list(paired=TRUE), map_signif_level = TRUE, 
                          textsize = 4) +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) +
    scale_y_continuous("ARI", breaks=c(0,0.25, 0.5, 0.75, 1), limits = c(NA, 1.3)) +
    scale_x_discrete("", labels=c(k_mode="k-mode", middir_dp="mixdir DP", mixdir="mixdir", poLCA="poLCA", rock="ROCK")) +
    ggtitle("True Classes = 4 and K=8")

p_comb <- cowplot::plot_grid(p1, p2, align = "v", labels=c("A", "B"))
p_comb
cowplot::ggsave("../output/performance_comparison.png", p_comb, width=7, height=4.5)
```



```{r}

p1 <- data %>%
  dplyr::select(-arand_idx) %>%
  filter(method != "correct" & method != "all_one") %>%
  mutate(method= forcats::fct_reorder(as.factor(method), time, mean)) %>%
  ggplot(aes(x=method, y=time)) +
    stat_summary(fun.data = "mean_cl_boot", colour = "red", geom = "crossbar", size=0.2, width=0.5) +
    ggbeeswarm::geom_quasirandom(width = 0.2) +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) +
    scale_x_discrete("", labels=c(k_mode="k-mode", middir_dp="mixdir DP", mixdir="mixdir", poLCA="poLCA", rock="ROCK")) +
    ylab("Time [seconds]") +
    ggtitle("True Classes = 4 and K=4")

p2 <- data2 %>%
  dplyr::select(-arand_idx) %>%
  filter(method != "correct" & method != "all_one") %>%
  mutate(method= forcats::fct_reorder(as.factor(method), time, mean)) %>%
  ggplot(aes(x=method, y=time)) +
    stat_summary(fun.data = "mean_cl_boot", colour = "red", geom = "crossbar", size=0.2, width=0.5) +
    ggbeeswarm::geom_quasirandom(width = 0.2) +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) +
    # scale_y_continuous("ARI", breaks=c(0,0.25, 0.5, 0.75, 1), limits = c(NA, 1.5)) +
    scale_x_discrete("", labels=c(k_mode="k-mode", middir_dp="mixdir DP", mixdir="mixdir", poLCA="poLCA", rock="ROCK")) +
    ylab("Time [seconds]") +
    ggtitle("True Classes = 4 and K=8")

p_comb <- cowplot::plot_grid(p1, p2, align = "v", labels=c("A", "B"))
p_comb
cowplot::ggsave("../output/performance_comparison_timing.png", p_comb, width=7, height=4.5)
```


```{r}
run_mixdirdp_on_generated_data <- function(n_iter, n_latent){
  set.seed(1)
  lapply(1:n_iter, function(idx){
    
    tmp <- generate_categorical_dataset(n_ind=1000, n_quest=10, n_cat=rbinom(1, 30, 0.3), n_true_classes = 4, 
                                        lambda_true=extraDistr::rdirichlet(1, rep(1,4)),
                                        bb_size=100, bb_suc=0.2, bb_overdis=0.1)
    X <- tmp$X
    true_classes <- tmp$true_latent
    
    mixdirdp_result <- mixdir(X, n_latent=n_latent, select_latent = TRUE,
                              alpha=c(1,1), beta = 0.1, verbose=TRUE, max_iter = 300)
    
    list(true_classes=true_classes, mixdirdp_result=mixdirdp_result)
    
  })
}





data2 <- run_mixdirdp_on_generated_data(n_latent=8, n_iter=25)






data2 %>%
  keep(~ .$mixdirdp_result$converged) %>%
  map_dbl(~ mcclust::arandi(.$true_classes, .$mixdirdp_result$pred_class))
  

data2 %>%
  keep(~ .$mixdirdp_result$converged) %>%
  {left_join(imap_dfr(., ~ enframe(table(.x$mixdirdp_result$pred_class)) %>% 
                        mutate(idx=.y, 
                               ARI=mcclust::arandi(.x$true_classes, .x$mixdirdp_result$pred_class))),
             imap_dfr(., ~ enframe(table(.x$true_classes)) %>% mutate(idx=.y)),
             by=c("idx", "name")
            )} %>%
  filter(ARI > 0.5) %>%
  mutate(value.y=ifelse(is.na(value.y), 0, value.y)) %>%
  group_by(idx) %>%
  mutate(value.y=sort(value.y, decreasing = TRUE)) %>%
  gather(origin, value, c(value.x, value.y)) %>%
  mutate(origin=if_else(origin == "value.x", "Inferred", "Original")) %>%
  ungroup() %>%
  mutate(label=forcats::fct_reorder(as.factor(paste0("ARI: ", formatC(ARI, digits=3))), -ARI)) %>%
  ggplot(aes(x=name, y=value, fill=origin)) +
    geom_bar(stat="identity", position = "dodge") +
    facet_wrap(~ label)

```


