---
title: "Young Lives (Preparation)"
output: html_notebook
---

In this notebook I load and clean the Young Lives survey data to bring it in an appropriate shape for clustering.
Due to privacy reasons we cannot publicly share the raw data, but you can apply for access yourself [here]( https://discover.ukdataservice.ac.uk/catalogue?sn=7483) to reproduce all the preparation steps. After cleaning
the data we cluster the data using different hyperparameters. Because the result contains only aggregated data
we can share the resulting mixdir-objects (`data/young_lives_clustering_results`).

```{r}
library(tidyverse)
library(purrr)
library(stringr)
library(forcats)
```



# Check if data is accessible

```{r}
if(file.exists("../data/young_lives_data/UKDA-7483-tab")){
  print("Data available, reproducing data preparation and clustering")
}else{
  stop(paste0("Due to data privacy we cannot share this dataset. Please get if from ",
             "here (https://discover.ukdataservice.ac.uk/catalogue?sn=7483) to fully reproduce the analysis"))
}
```


# Load Data

# Ethiopia

Start with the analysis of Ethiopia:

```{r}
data_raw_ethiopia <- read_tsv("../data/young_lives_data/UKDA-7483-tab/tab/ethiopia_constructed.tab")

data_cleaned_ethiopia <- data_raw_ethiopia %>%
  # Remove some obviously useless variables
  dplyr::select(- c(dint, headid, careid, dadid, momid, clustid, commid)) %>%
  # Remove variables that are redundant
  dplyr::select( -c(totalexp, totalexp_pa, totalexp_rpa,
                    foodexp, foodexp_pa, foodexp_rpa,
                    nfoodexp, nfoodexp_pa, nfoodexp_rpa)) %>%
  # Remove variables that seem overly specific
  dplyr::select(- c(anidrau, anirumi, anispec, anicowm, anicowt,
                    anicalv, anibufm,anibuft, aniheif, anibull, anihebu, 
                    anidonk, aniybul, anioxen, anicaml, anishee, anigoat, 
                    anipigs, anipoul, anirabb, anibeeh, aniothr)) %>%
  dplyr::select(- c(male05, male612, male1317, male1860, male61,
                    female05, female612, female1317, female1860, female61)) %>%
  dplyr::select(- c(fhfa, fwfa, fbfa, fwfl)) %>%
  # A lot of variables are on a scale, I will cut them
  mutate(headage=cut(headage, breaks=5),
         headedu=case_when(
           headedu < 4 ~ "<4",
           headedu < 7 ~ ">=4&<7",
           headedu < 10 ~ ">=7&<10",
           headedu < 14 ~ ">=10&<14",
           headedu == 28 ~ "adult literacy",
           headedu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         wi=cut(wi, breaks=5),
         hq=cut(hq, breaks=5),
         cd=cut(cd, breaks=5),
         foodexp_r=cut(foodexp_r, breaks=5),
         nfoodexp_r=cut(nfoodexp_r, breaks=5),
         totalexp_r=cut(totalexp_r, breaks=5),
         animilk=fct_lump(as.factor(animilk), n=5, other_level = "more"),
         ownland=cut(log(ownland + 0.001), breaks=5),
         totland=cut(log(totland + 0.001), breaks=5),
         hhsize=cut(hhsize, breaks=5),
         careage=cut(careage, breaks=5),
         caredu=case_when(
           caredu < 4 ~ "<4",
           caredu < 7 ~ ">=4&<7",
           caredu < 10 ~ ">=7&<10",
           caredu < 14 ~ ">=10&<14",
           caredu == 28 ~ "adult literacy",
           caredu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         agemon=cut(agemon, breaks=4),
         chweight=cut(chweight, breaks=5),
         chheight=cut(chheight, breaks=5),
         bwght=cut(bwght, breaks=5),
         bmi=cut(bmi, breaks=5),
         zwfa=cut(zwfa, breaks=5),
         zhfa=cut(zhfa, breaks=5),
         zbfa=cut(zbfa, breaks=5),
         zwfl=cut(zwfl, breaks=5),
         dadage=cut(dadage, breaks=5),
         dadedu=case_when(
           dadedu < 4 ~ "<4",
           dadedu < 7 ~ ">=4&<7",
           dadedu < 10 ~ ">=7&<10",
           dadedu < 14 ~ ">=10&<14",
           dadedu == 28 ~ "adult literacy",
           dadedu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         momage=cut(momage, breaks=5),
         momedu=case_when(
           momedu < 4 ~ "<4",
           momedu < 7 ~ ">=4&<7",
           momedu < 10 ~ ">=7&<10",
           momedu < 14 ~ ">=10&<14",
           momedu == 28 ~ "adult literacy",
           momedu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         timesch=cut(timesch, breaks=5),
         chgrade=case_when(
           chgrade < 4 ~ "<4",
           chgrade < 7 ~ ">=4&<7",
           chgrade < 10 ~ ">=7&<10",
           chgrade < 14 ~ ">=10&<14",
           chgrade == 28 ~ "adult literacy",
           chgrade == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         ravens=cut(ravens, breaks=5),
         ppvtraw=cut(ppvtraw, breaks=5),
         hsleep=cut(hsleep, breaks=5),
         hcare=cut(hcare, breaks=5),
         hchore=cut(hchore, breaks=5),
         htask=cut(htask, breaks=5),
         hwork=cut(hwork, breaks=5),
         hschool=cut(hschool, breaks=5),
         hstudy=cut(hstudy, breaks=5),
         hplay=cut(hplay, breaks=5)
       )  %>%
    # Drop columns with only 1 value
    select_if(~ length(unique(.x)) >= 2)

```




```{r}
young_cohort_all_points_ethiopia <- data_cleaned_ethiopia %>%
  filter(yc == 1 & panel1234 == 1)

first_rounder_ethiopia <- young_cohort_all_points_ethiopia %>%
  filter(round == 1) %>%
  select_if(~ length(unique(.x)) >= 2)

second_rounder_ethiopia <- young_cohort_all_points_ethiopia %>%
  filter(round == 2) %>%
  select_if(~ length(unique(.x)) >= 2)

third_rounder_ethiopia <- young_cohort_all_points_ethiopia %>%
  filter(round == 3) %>%
  select_if(~ length(unique(.x)) >= 2)

fourth_rounder_ethiopia <- young_cohort_all_points_ethiopia %>%
  filter(round == 4) %>%
  select_if(~ length(unique(.x)) >= 2)

common_colnames_ethiopia <- intersect(intersect(colnames(first_rounder_ethiopia), colnames(second_rounder_ethiopia)),
          intersect(colnames(third_rounder_ethiopia), colnames(fourth_rounder_ethiopia)))

first_rounder_ethiopia <- first_rounder_ethiopia[, common_colnames_ethiopia]
second_rounder_ethiopia <- second_rounder_ethiopia[, common_colnames_ethiopia]
third_rounder_ethiopia <- third_rounder_ethiopia[, common_colnames_ethiopia]
fourth_rounder_ethiopia <- fourth_rounder_ethiopia[, common_colnames_ethiopia]


```







# India

```{r}
data_raw_india <- read_tsv("../data/young_lives_data/young_lives_data/UKDA-7483-tab/tab/india_constructed.tab")

data_cleaned_india <- data_raw_india %>%
  # Remove some obviously useless variables
  dplyr::select(- c(dint, headid, careid, dadid, momid, clustid, commid)) %>%
  # Remove variables that are redundant
  dplyr::select( -c(totalexp, totalexp_pc, totalexp_rpc,
                    foodexp, foodexp_pc, foodexp_rpc,
                    nfoodexp, nfoodexp_pc, nfoodexp_rpc)) %>%
  # Remove variables that seem overly specific
  dplyr::select(- c(anidrau, anirumi, anispec, anicowm, anicowt,
                    anicalv, anibufm,anibuft, aniheif, anibull, anihebu, 
                    anidonk, aniybul, anishee, anigoat, 
                    anifish,anishri,anifshr,
                    anipigs, anipoul, anirabb, anibeeh, aniothr)) %>%
  dplyr::select(- c(male05, male612, male1317, male1860, male61,
                    female05, female612, female1317, female1860, female61)) %>%
  dplyr::select(- c(fhfa, fwfa, fbfa, fwfl))  %>%
  # A lot of variables are on a scale, I will cut them
  mutate(headage=cut(headage, breaks=5),
         headedu=case_when(
           headedu < 4 ~ "<4",
           headedu < 7 ~ ">=4&<7",
           headedu < 10 ~ ">=7&<10",
           headedu < 14 ~ ">=10&<14",
           headedu == 28 ~ "adult literacy",
           headedu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         wi=cut(wi, breaks=5),
         hq=cut(hq, breaks=5),
         cd=cut(cd, breaks=5),
         foodexp_r=cut(foodexp_r, breaks=5),
         nfoodexp_r=cut(nfoodexp_r, breaks=5),
         totalexp_r=cut(totalexp_r, breaks=5),
         animilk=fct_lump(as.factor(animilk), n=5, other_level = "more"),
         ownland=cut(log(ownland + 0.001), breaks=5),
         totland=cut(log(totland + 0.001), breaks=5),
         hhsize=cut(hhsize, breaks=5),
         careage=cut(careage, breaks=5),
         caredu=case_when(
           caredu < 4 ~ "<4",
           caredu < 7 ~ ">=4&<7",
           caredu < 10 ~ ">=7&<10",
           caredu < 14 ~ ">=10&<14",
           caredu == 28 ~ "adult literacy",
           caredu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         agemon=cut(agemon, breaks=4),
         chweight=cut(chweight, breaks=5),
         chheight=cut(chheight, breaks=5),
         bwght=cut(bwght, breaks=5),
         bmi=cut(bmi, breaks=5),
         zwfa=cut(zwfa, breaks=5),
         zhfa=cut(zhfa, breaks=5),
         zbfa=cut(zbfa, breaks=5),
         zwfl=cut(zwfl, breaks=5),
         dadage=cut(dadage, breaks=5),
         dadedu=case_when(
           dadedu < 4 ~ "<4",
           dadedu < 7 ~ ">=4&<7",
           dadedu < 10 ~ ">=7&<10",
           dadedu < 14 ~ ">=10&<14",
           dadedu == 28 ~ "adult literacy",
           dadedu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         momage=cut(momage, breaks=5),
         momedu=case_when(
           momedu < 4 ~ "<4",
           momedu < 7 ~ ">=4&<7",
           momedu < 10 ~ ">=7&<10",
           momedu < 14 ~ ">=10&<14",
           momedu == 28 ~ "adult literacy",
           momedu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         timesch=cut(timesch, breaks=5),
         chgrade=case_when(
           chgrade < 4 ~ "<4",
           chgrade < 7 ~ ">=4&<7",
           chgrade < 10 ~ ">=7&<10",
           chgrade < 14 ~ ">=10&<14",
           chgrade == 28 ~ "adult literacy",
           chgrade == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         ravens=cut(ravens, breaks=5),
         ppvtraw=cut(ppvtraw, breaks=5),
         hsleep=cut(hsleep, breaks=5),
         hcare=cut(hcare, breaks=5),
         hchore=cut(hchore, breaks=5),
         htask=cut(htask, breaks=5),
         hwork=cut(hwork, breaks=5),
         hschool=cut(hschool, breaks=5),
         hstudy=cut(hstudy, breaks=5),
         hplay=cut(hplay, breaks=5)
       )  %>%
    # Drop columns with only 1 value
    select_if(~ length(unique(.x)) >= 2)


```

```{r}
young_cohort_all_points_india <- data_cleaned_india %>%
  filter(yc == 1 & panel1234 == 1)

first_rounder_india <- young_cohort_all_points_india %>%
  filter(round == 1) %>%
  select_if(~ length(unique(.x)) >= 2)

second_rounder_india <- young_cohort_all_points_india %>%
  filter(round == 2) %>%
  select_if(~ length(unique(.x)) >= 2)

third_rounder_india <- young_cohort_all_points_india %>%
  filter(round == 3) %>%
  select_if(~ length(unique(.x)) >= 2)

fourth_rounder_india <- young_cohort_all_points_india %>%
  filter(round == 4) %>%
  select_if(~ length(unique(.x)) >= 2)

common_colnames_india <- intersect(intersect(colnames(first_rounder_india), colnames(second_rounder_india)),
          intersect(colnames(third_rounder_india), colnames(fourth_rounder_india)))

first_rounder_india <- first_rounder_india[, common_colnames_india]
second_rounder_india <- second_rounder_india[, common_colnames_india]
third_rounder_india <- third_rounder_india[, common_colnames_india]
fourth_rounder_india <- fourth_rounder_india[, common_colnames_india]

```




# Peru 


```{r}
data_raw_peru <- read_tsv("../data/young_lives_data/UKDA-7483-tab/tab/peru_constructed.tab")

data_cleaned_peru <- data_raw_peru %>%
  # Remove some obviously useless variables
  dplyr::select(- c(dint, headid, careid, dadid, momid, clustid, placeid)) %>%
  # Remove variables that are redundant
  dplyr::select( -c(totalexp, totalexp_pc, totalexp_rpc,
                    foodexp, foodexp_pc, foodexp_rpc,
                    nfoodexp, nfoodexp_pc, nfoodexp_rpc))  %>%
  # Remove variables that seem overly specific
  dplyr::select(- c(anidrau, anirumi, anispec, anicowm, anicowt,
                    anidonk, anishee, anigoat, 
                    anioxen,anillam,aniguin,anisnai,
                    anifish,anishri,anifshr,
                    anipigs, anipoul, anirabb, anibeeh, aniothr)) %>%
  dplyr::select(- c(male05, male612, male1317, male1860, male61,
                    female05, female612, female1317, female1860, female61)) %>%
  dplyr::select(- c(fhfa, fwfa, fbfa, fwfl))  %>%
  # A lot of variables are on a scale, I will cut them
  mutate(headage=cut(headage, breaks=5),
         headedu=case_when(
           headedu < 4 ~ "<4",
           headedu < 7 ~ ">=4&<7",
           headedu < 10 ~ ">=7&<10",
           headedu < 14 ~ ">=10&<14",
           headedu == 28 ~ "adult literacy",
           headedu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         wi=cut(wi, breaks=5),
         hq=cut(hq, breaks=5),
         cd=cut(cd, breaks=5),
         foodexp_r=cut(foodexp_r, breaks=5),
         nfoodexp_r=cut(nfoodexp_r, breaks=5),
         totalexp_r=cut(totalexp_r, breaks=5),
         animilk=fct_lump(as.factor(animilk), n=5, other_level = "more"),
         ownland=cut(log(ownland + 0.001), breaks=5),
         totland=cut(log(totland + 0.001), breaks=5),
         hhsize=cut(hhsize, breaks=5),
         careage=cut(careage, breaks=5),
         caredu=case_when(
           caredu < 4 ~ "<4",
           caredu < 7 ~ ">=4&<7",
           caredu < 10 ~ ">=7&<10",
           caredu < 14 ~ ">=10&<14",
           caredu == 28 ~ "adult literacy",
           caredu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         agemon=cut(agemon, breaks=4),
         chweight=cut(chweight, breaks=5),
         chheight=cut(chheight, breaks=5),
         bwght=cut(bwght, breaks=5),
         bmi=cut(bmi, breaks=5),
         zwfa=cut(zwfa, breaks=5),
         zhfa=cut(zhfa, breaks=5),
         zbfa=cut(zbfa, breaks=5),
         zwfl=cut(zwfl, breaks=5),
         dadage=cut(dadage, breaks=5),
         dadedu=case_when(
           dadedu < 4 ~ "<4",
           dadedu < 7 ~ ">=4&<7",
           dadedu < 10 ~ ">=7&<10",
           dadedu < 14 ~ ">=10&<14",
           dadedu == 28 ~ "adult literacy",
           dadedu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         momage=cut(momage, breaks=5),
         momedu=case_when(
           momedu < 4 ~ "<4",
           momedu < 7 ~ ">=4&<7",
           momedu < 10 ~ ">=7&<10",
           momedu < 14 ~ ">=10&<14",
           momedu == 28 ~ "adult literacy",
           momedu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         timesch=cut(timesch, breaks=5),
         chgrade=case_when(
           chgrade < 4 ~ "<4",
           chgrade < 7 ~ ">=4&<7",
           chgrade < 10 ~ ">=7&<10",
           chgrade < 14 ~ ">=10&<14",
           chgrade == 28 ~ "adult literacy",
           chgrade == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         ravens=cut(ravens, breaks=5),
         ppvtraw=cut(ppvtraw, breaks=5),
         hsleep=cut(hsleep, breaks=5),
         hcare=cut(hcare, breaks=5),
         hchore=cut(hchore, breaks=5),
         htask=cut(htask, breaks=5),
         hwork=cut(hwork, breaks=5),
         hschool=cut(hschool, breaks=5),
         hstudy=cut(hstudy, breaks=5),
         hplay=cut(hplay, breaks=5)
       )  %>%
    # Drop columns with only 1 value
    select_if(~ length(unique(.x)) >= 2)


```


```{r}
young_cohort_all_points_peru <- data_cleaned_peru %>%
  filter(yc == 1 & panel1234 == 1)

first_rounder_peru <- young_cohort_all_points_peru %>%
  filter(round == 1) %>%
  select_if(~ length(unique(.x)) >= 2)

second_rounder_peru <- young_cohort_all_points_peru %>%
  filter(round == 2) %>%
  select_if(~ length(unique(.x)) >= 2)

third_rounder_peru <- young_cohort_all_points_peru %>%
  filter(round == 3) %>%
  select_if(~ length(unique(.x)) >= 2)

fourth_rounder_peru <- young_cohort_all_points_peru %>%
  filter(round == 4) %>%
  select_if(~ length(unique(.x)) >= 2)

common_colnames_peru <- intersect(intersect(colnames(first_rounder_peru), colnames(second_rounder_peru)),
          intersect(colnames(third_rounder_peru), colnames(fourth_rounder_peru)))

first_rounder_peru <- first_rounder_peru[, common_colnames_peru]
second_rounder_peru <- second_rounder_peru[, common_colnames_peru]
third_rounder_peru <- third_rounder_peru[, common_colnames_peru]
fourth_rounder_peru <- fourth_rounder_peru[, common_colnames_peru]


```










# Vietnam 



```{r}
data_raw_vietnam <- read_tsv("../data/young_lives_data/UKDA-7483-tab/tab/vietnam_constructed.tab")

data_cleaned_vietnam <- data_raw_vietnam %>%
  # Remove some obviously useless variables
  dplyr::select(- c(dint, headid, careid, dadid, momid, clustid, commid)) %>%
  # Remove variables that are redundant
  dplyr::select( -c(totalexp, totalexp_pc, totalexp_rpc,
                    foodexp, foodexp_pc, foodexp_rpc,
                    nfoodexp, nfoodexp_pc, nfoodexp_rpc, 
                    elecqq, cookingqq,drwaterqq, toiletqq)) %>%
  # Remove variables that seem overly specific
  dplyr::select(- c(anidrau, anirumi, anispec, anicowm, anicowt,
                    anicalv, anibufm,anibuft, anibull, anihebu, 
                    anidonk, anishee, anigoat, 
                    anipigs, anipoul, anirabb, aniothr)) %>%
  dplyr::select(- c(male05, male612, male1317, male1860, male61,
                    female05, female612, female1317, female1860, female61)) %>%
  dplyr::select(- c(fhfa, fwfa, fbfa, fwfl))  %>%
  # A lot of variables are on a scale, I will cut them
  mutate(headage=cut(headage, breaks=5),
         headedu=case_when(
           headedu < 4 ~ "<4",
           headedu < 7 ~ ">=4&<7",
           headedu < 10 ~ ">=7&<10",
           headedu < 14 ~ ">=10&<14",
           headedu == 28 ~ "adult literacy",
           headedu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         wi=cut(wi, breaks=5),
         hq=cut(hq, breaks=5),
         cd=cut(cd, breaks=5),
         foodexp_r=cut(foodexp_r, breaks=5),
         nfoodexp_r=cut(nfoodexp_r, breaks=5),
         totalexp_r=cut(totalexp_r, breaks=5),
         animilk=fct_lump(as.factor(animilk), n=5, other_level = "more"),
         ownland=cut(log(ownland + 0.001), breaks=5),
         totland=cut(log(totland + 0.001), breaks=5),
         hhsize=cut(hhsize, breaks=5),
         careage=cut(careage, breaks=5),
         caredu=case_when(
           caredu < 4 ~ "<4",
           caredu < 7 ~ ">=4&<7",
           caredu < 10 ~ ">=7&<10",
           caredu < 14 ~ ">=10&<14",
           caredu == 28 ~ "adult literacy",
           caredu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         agemon=cut(agemon, breaks=4),
         chweight=cut(chweight, breaks=5),
         chheight=cut(chheight, breaks=5),
         bwght=cut(bwght, breaks=5),
         bmi=cut(bmi, breaks=5),
         zwfa=cut(zwfa, breaks=5),
         zhfa=cut(zhfa, breaks=5),
         zbfa=cut(zbfa, breaks=5),
         zwfl=cut(zwfl, breaks=5),
         dadage=cut(dadage, breaks=5),
         dadedu=case_when(
           dadedu < 4 ~ "<4",
           dadedu < 7 ~ ">=4&<7",
           dadedu < 10 ~ ">=7&<10",
           dadedu < 14 ~ ">=10&<14",
           dadedu == 28 ~ "adult literacy",
           dadedu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         momage=cut(momage, breaks=5),
         momedu=case_when(
           momedu < 4 ~ "<4",
           momedu < 7 ~ ">=4&<7",
           momedu < 10 ~ ">=7&<10",
           momedu < 14 ~ ">=10&<14",
           momedu == 28 ~ "adult literacy",
           momedu == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         timesch=cut(timesch, breaks=5),
         chgrade=case_when(
           chgrade < 4 ~ "<4",
           chgrade < 7 ~ ">=4&<7",
           chgrade < 10 ~ ">=7&<10",
           chgrade < 14 ~ ">=10&<14",
           chgrade == 28 ~ "adult literacy",
           chgrade == 29 ~ "religous edu",
           TRUE ~ "other"
         ),
         ravens=cut(ravens, breaks=5),
         ppvtraw=cut(ppvtraw, breaks=5),
         hsleep=cut(hsleep, breaks=5),
         hcare=cut(hcare, breaks=5),
         hchore=cut(hchore, breaks=5),
         htask=cut(htask, breaks=5),
         hwork=cut(hwork, breaks=5),
         hschool=cut(hschool, breaks=5),
         hstudy=cut(hstudy, breaks=5),
         hplay=cut(hplay, breaks=5)
       )  %>%
    # Drop columns with only 1 value
    select_if(~ length(unique(.x)) >= 2)

```




```{r}
young_cohort_all_points_vietnam <- data_cleaned_vietnam %>%
  filter(yc == 1 & panel1234 == 1)

first_rounder_vietnam <- young_cohort_all_points_vietnam %>%
  filter(round == 1) %>%
  select_if(~ length(unique(.x)) >= 2)

second_rounder_vietnam <- young_cohort_all_points_vietnam %>%
  filter(round == 2) %>%
  select_if(~ length(unique(.x)) >= 2)

third_rounder_vietnam <- young_cohort_all_points_vietnam %>%
  filter(round == 3) %>%
  select_if(~ length(unique(.x)) >= 2)

fourth_rounder_vietnam <- young_cohort_all_points_vietnam %>%
  filter(round == 4) %>%
  select_if(~ length(unique(.x)) >= 2)

common_colnames_vietnam <- intersect(intersect(colnames(first_rounder_vietnam), colnames(second_rounder_vietnam)),
          intersect(colnames(third_rounder_vietnam), colnames(fourth_rounder_vietnam)))

first_rounder_vietnam <- first_rounder_vietnam[, common_colnames_vietnam]
second_rounder_vietnam <- second_rounder_vietnam[, common_colnames_vietnam]
third_rounder_vietnam <- third_rounder_vietnam[, common_colnames_vietnam]
fourth_rounder_vietnam <- fourth_rounder_vietnam[, common_colnames_vietnam]

```











# Performance comparison

```{r}
library(mixdir)
library(poLCA)
library(klaR)
library(cba)
library(parallel)
```



```{r}
test_cluster_perf <- function(n_iter, n_latent,
                              first_rounder, second_rounder, third_rounder, fourth_rounder){

  library(poLCA)

  # parallel::mclapply(1:n_iter, mc.cores=min(4,n_iter), function(idx){
  lapply(1:n_iter, function(idx){
    set.seed(idx)
    selection <- sample(1:nrow(first_rounder), size=round(0.2 * nrow(first_rounder)))
    X1 <- map_df(first_rounder[selection, ], ~ as.numeric(addNA(as.factor(.x))))[, -1]
    X2 <- map_df(second_rounder[selection, ], ~ as.numeric(addNA(as.factor(.x))))[, -1]
    X3 <- map_df(third_rounder[selection, ], ~ as.numeric(addNA(as.factor(.x))))[, -1]
    X4 <- map_df(fourth_rounder[selection, ], ~ as.numeric(addNA(as.factor(.x))))[, -1]
    all_X <- list(X1, X2, X3, X4)
    print(paste0("Starting iter", idx))



    mixdir_results <- lapply(all_X, function(X){
      timing <- system.time(mixdir_result <- mixdir(X, n_latent=n_latent, select_latent = FALSE, alpha=1, beta = 0.1))
      list(clustering=mixdir_result, timing=timing)
    })

    print("Finised Mixdir")

    mixdirdp_results <- lapply(all_X, function(X){
      timing <- system.time(mixdir_result <- mixdir(X, n_latent=n_latent, select_latent = TRUE, alpha=c(1,1), beta = 0.1))
      list(clustering=mixdir_result, timing=timing)
    })


    print("Finished MixdirDP")

     mixdirdp100_results <- lapply(all_X, function(X){
      timing <- system.time(mixdir_result <- mixdir(X, n_latent=n_latent, select_latent = TRUE, alpha=c(100,1), beta = 0.1))
      list(clustering=mixdir_result, timing=timing)
    })

    print("Finished MixdirDP100")

    kmode_results <- lapply(all_X, function(X){
      timing <- system.time(kmode_result <- klaR::kmodes(X,n_latent))
      list(clustering=kmode_result, timing=timing)
    })


    print("Finished kmode")

    rock_results <- lapply(all_X, function(X){
      timing <- system.time(rock_result <- fitted(cba::rockCluster(cba::as.dummy(map_df(X, ~ as.integer(as.factor(.x)))),
                                                                    n=n_latent, theta = 0.1, debug=FALSE)))
      list(clustering=rock_result, timing=timing)
    })

    print("Finished ROCK")

    lca_results <- lapply(all_X, function(X){
      timing <- system.time(lca_result <- poLCA(formula(paste0("cbind(", paste0(colnames(X1), collapse = ","), ") ~ 1")),
                        X, nclass=n_latent, verbose = FALSE))
      list(clustering=lca_result, timing=timing)
    })

    print("Finished poLCA")

    list(mixdir_results=mixdir_results,
         mixdirdp_results=mixdirdp_results,
         mixdirdp100_results=mixdirdp100_results,
         kmode_results=kmode_results,
         rock_results=rock_results,
         lca_results=lca_results)

  })
}


```



Run Clustering

**Warning:** This might take quite a while to complete, so it is a good idea to run it on a cluster in parrallel

```{r}
# Ethiopia 
result_ethiopia <- test_cluster_perf(n_iter=5, n_latent=5,  
                                     first_rounder_ethiopia,
                                     second_rounder_ethiopia,
                                     third_rounder_ethiopia,
                                     fourth_rounder_ethiopia)
saveRDS(result_ethiopia, "../data/young_lives_clustering_results/young_survey_perf_comp_ethiopia_lat5.RDS")

result_ethiopia <- test_cluster_perf(n_iter=25, n_latent=25,  
                                     first_rounder_ethiopia,
                                     second_rounder_ethiopia,
                                     third_rounder_ethiopia,
                                     fourth_rounder_ethiopia)
saveRDS(result_ethiopia, "../data/young_lives_clustering_results/young_survey_perf_comp_ethiopia_lat25.RDS")


# India
result_india <- test_cluster_perf(n_iter=5, n_latent=5,  
                                     first_rounder_india,
                                     second_rounder_india,
                                     third_rounder_india,
                                     fourth_rounder_india)
saveRDS(result_india, "../data/young_lives_clustering_results/young_survey_perf_comp_india_lat5.RDS")

result_india <- test_cluster_perf(n_iter=25, n_latent=25,  
                                     first_rounder_india,
                                     second_rounder_india,
                                     third_rounder_india,
                                     fourth_rounder_india)
saveRDS(result_india, "../data/young_lives_clustering_results/young_survey_perf_comp_india_lat25.RDS")

# Peru
result_peru <- test_cluster_perf(n_iter=5, n_latent=5,  
                                     first_rounder_peru,
                                     second_rounder_peru,
                                     third_rounder_peru,
                                     fourth_rounder_peru)
saveRDS(result_peru, "../data/young_lives_clustering_results/young_survey_perf_comp_peru_lat5.RDS")

result_peru <- test_cluster_perf(n_iter=25, n_latent=25,  
                                     first_rounder_peru,
                                     second_rounder_peru,
                                     third_rounder_peru,
                                     fourth_rounder_peru)
saveRDS(result_peru, "../data/young_lives_clustering_results/young_survey_perf_comp_peru_lat25.RDS")

# Vietnam
result_vietnam <- test_cluster_perf(n_iter=5, n_latent=5,  
                                     first_rounder_vietnam,
                                     second_rounder_vietnam,
                                     third_rounder_vietnam,
                                     fourth_rounder_vietnam)

saveRDS(result_vietnam, "../data/young_lives_clustering_results/young_survey_perf_comp_vietnam_lat5.RDS")

result_vietnam <- test_cluster_perf(n_iter=25, n_latent=25,  
                                     first_rounder_vietnam,
                                     second_rounder_vietnam,
                                     third_rounder_vietnam,
                                     fourth_rounder_vietnam)

saveRDS(result_vietnam, "../data/young_lives_clustering_results/young_survey_perf_comp_vietnam_lat25.RDS")

```






# Prepare data for alluvial plot

```{r}
set.seed(1)
res1 <- mixdir(first_rounder_ethiopia[, -1], n_latent=5, select_latent = FALSE, beta=5,
               max_iter=200, repetitions = 5, verbose=TRUE)
res2 <- mixdir(first_rounder_ethiopia[, -1], n_latent=5, select_latent = FALSE, beta=5,
               max_iter=200, repetitions = 5, verbose=TRUE)
res3 <- mixdir(first_rounder_ethiopia[, -1], n_latent=5, select_latent = FALSE, beta=5,
               max_iter=200, repetitions = 5, verbose=TRUE)
res4 <- mixdir(first_rounder_ethiopia[, -1], n_latent=5, select_latent = FALSE, beta=5,
               max_iter=200, repetitions = 5, verbose=TRUE)

set.seed(1)
res1_rep <- lapply(1:5, function(x){
  mixdir(first_rounder_ethiopia[, -1], n_latent=5, select_latent = FALSE, beta=5,
               max_iter=200, repetitions = 1, verbose=TRUE)
})

write_rds(list(res1, res2, res3, res4, res1_rep), "../data/young_lives_clustering_results/young_lives_res1234_resrep.RDS")
```













