---
title: "Young Lives (Analysis)"
output: html_notebook
---

In this notebook I evaluate the clustering results on the Young Lives for all four locations India, Vietnam, Peru and Ethiopia.
The young lives survey follows children across 4 time points which allows us to evaluate the clustering quality using the 
consistency of the clusterings over time. A good method should be able usually to assign a child to the same cluster regardless of the
particular time point.


```{r setup}
library(tidyverse)
library(purrr)
library(stringr)
library(forcats)
library(mixdir)
library(cowplot)
set.seed(1)
```





# Check performance on all 4 sub-studies (Fig. 8)

Load clustering results for India, Vietnam, Peru and Ethiopia

```{r}

perf_comparison_countries <- list(India=read_rds("../data/young_lives_clustering_results/young_survey_perf_comp_india_lat5.RDS"),
                                  Vietnam=read_rds("../data/young_lives_clustering_results/young_survey_perf_comp_vietnam_lat5.RDS"),
                                  Peru=read_rds("../data/young_lives_clustering_results/young_survey_perf_comp_peru_lat5.RDS"),
                                  Ethiopia=read_rds("../data/young_lives_clustering_results/young_survey_perf_comp_ethiopia_lat5.RDS"))

class_assignment <- map2_df(perf_comparison_countries, names(perf_comparison_countries), function(perf_comparison, country){
  map2_df(perf_comparison, seq_along(perf_comparison), function(perf, iter){
    map2_df(perf, names(perf), function(res_list, name){
      method <- str_split(name, "_")[[1]][1]
      map2_df(res_list, seq_along(res_list), function(res_cont, round){
        res <- res_cont$clustering
        classes <- if(method == "mixdir" | method == "mixdirdp" | method == "mixdirdp100"){
          res$pred_class
        }else if(method == "lca"){
          res$predclass
        }else if(method == "kmode"){
          res$cluster
        }else if(method == "rock"){
          as.integer(res$cl)
        }else{
          stop("Should not come here")
        }
        data_frame(country=country, iter=iter, method=method, round=round, class=classes, ind=1:length(classes), time=res_cont$timing["user.self"])
      })
    })
  })
})

perf_comparison_countries25 <- list(India=read_rds("../data/young_lives_clustering_results/young_survey_perf_comp_india_lat25.RDS"),
                                    Vietnam=read_rds("../data/young_lives_clustering_results/young_survey_perf_comp_vietnam_lat25.RDS"),
                                    Peru=read_rds("../data/young_lives_clustering_results/young_survey_perf_comp_peru_lat25.RDS"),
                                    Ethiopia=read_rds("../data/young_lives_clustering_results/young_survey_perf_comp_ethiopia_lat25.RDS"))

class_assignment25 <- map2_df(perf_comparison_countries25, names(perf_comparison_countries25), function(perf_comparison, country){
  map2_df(perf_comparison, seq_along(perf_comparison), function(perf, iter){
    map2_df(perf, names(perf), function(res_list, name){
      method <- str_split(name, "_")[[1]][1]
      map2_df(res_list, seq_along(res_list), function(res_cont, round){
        res <- res_cont$clustering
        classes <- if(method == "mixdir" | method == "mixdirdp" | method == "mixdirdp100"){
          res$pred_class
        }else if(method == "lca"){
          res$predclass
        }else if(method == "kmode"){
          res$cluster
        }else if(method == "rock"){
          as.integer(res$cl)
        }else{
          stop("Should not come here")
        }
        data_frame(country=country, iter=iter, method=method, round=round, class=classes, ind=1:length(classes), time=res_cont$timing["user.self"])
      })
    })
  })
})

```


Create the clustering consistency beeswarm plot for K=5

```{r}
p_perf_lat5 <- class_assignment %>%
  filter(method != "mixdirdp100") %>%
  mutate(class = as.integer(as.factor(class))) %>%
  group_by(country, method, iter) %>%
  do({
    tmp <- .
    cross_df(list(a=unique(tmp$round), b=unique(tmp$round))) %>%
      filter(a < b) %>%
      rowwise() %>%
      mutate(ARI= mcclust::arandi(tmp[tmp$round == a, ]$class,
                                  tmp[tmp$round == b, ]$class))
  }) %>%
  ungroup() %>%
  mutate(method= fct_reorder(as.factor(method), -ARI, median)) %>%
  ggplot(aes(x=method, y=ARI)) + 
    ggbeeswarm::geom_quasirandom() +
    stat_summary(fun.data = "mean_cl_boot", colour = "red", geom = "crossbar", size=0.4) +
    # ggsignif::geom_signif(comparisons = list(c("lca", "mixdir"), c("mixdir", "mixdirdp"), c("mixdirdp", "kmode")),
    #                       y_position=c(0.82, 0.74, 0.6), map_signif_level = TRUE, test.args=list(paired=TRUE),
    #                       textsize = 4) +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) +
    facet_wrap(~ country) +
    scale_x_discrete("", labels=c(kmode="k-mode", mixdirdp="mixdir DP",mixdirdp100="mixdir DP\n(α₁=100)", mixdir="mixdir", lca="poLCA", rock="ROCK")) +
    scale_y_continuous(limits=c(NA, 0.75), breaks=c(0, 0.25, 0.5, 0.75, 1)) +
    ggtitle("Clustering Consistency (K=5)") 
```


Create the speed beeswarm plot for K=5

```{r}
p_dur_lat5 <- class_assignment  %>%
  filter(method != "mixdirdp100") %>%
  group_by(country, method, iter, round) %>%
  summarize(duration=first(time)) %>%
  ungroup() %>%
  mutate(method= fct_reorder(as.factor(method), duration, median)) %>%
  ggplot(aes(x=method, y=duration)) + 
    ggbeeswarm::geom_quasirandom() +
    stat_summary(fun.data = "mean_cl_boot", colour = "red", geom = "crossbar", size=0.4) +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) +
    scale_x_discrete("", labels=c(kmode="k-mode", mixdirdp100="mixdir DP\n(α₁=100)", mixdirdp="mixdir DP", mixdir="mixdir", lca="poLCA", rock="ROCK")) +
    scale_y_continuous(breaks=c(0.5, 1, 5, 10, 30, 60, 300, 600, 1800, 3600), minor_breaks = seq(1,10000, by=10),
                       labels=c("500 ms.", "1 s", "5 s", "10 s", "30 s", "1 min", "5 min", "10 min", "30 min", "1 h")) +
    facet_wrap(~ country) +
    ggtitle("Clustering Duration (K=5)") +
    coord_trans(y="log")
```


Create the clustering consistency beeswarm plot for K=25

```{r}
p_perf_lat25 <- class_assignment25 %>%
  filter(method != "mixdirdp100") %>%
  mutate(class = as.integer(as.factor(class))) %>%
  group_by(country, method, iter) %>%
  do({
    tmp <- .
    cross_df(list(a=unique(tmp$round), b=unique(tmp$round))) %>%
      filter(a < b) %>%
      rowwise() %>%
      mutate(ARI= mcclust::arandi(tmp[tmp$round == a, ]$class,
                                  tmp[tmp$round == b, ]$class))
  }) %>%
  ungroup() %>%
  mutate(method= fct_reorder(as.factor(method), -ARI, median)) %>%
  ggplot(aes(x=method, y=ARI)) + 
    ggbeeswarm::geom_quasirandom() +
    stat_summary(fun.data = "mean_cl_boot", colour = "red", geom = "crossbar", size=0.4) +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) +
    facet_wrap(~ country) +
    scale_x_discrete("", labels=c(kmode="k-mode", mixdirdp="mixdir DP",mixdirdp100="mixdir DP\n(α₁=100)", mixdir="mixdir", lca="poLCA", rock="ROCK")) +
    scale_y_continuous(limits=c(NA, 0.75), breaks=c(0, 0.25, 0.5, 0.75, 1)) +
    ggtitle("Clustering Consistency (K=25)") 
```

Create the speed beeswarm plot for K=5

```{r}
p_dur_lat25 <- class_assignment25  %>%
  filter(method != "mixdirdp100") %>%
  group_by(country, method, iter, round) %>%
  summarize(duration=first(time)) %>%
  ungroup() %>%
  mutate(method= fct_reorder(as.factor(method), duration, median)) %>%
  ggplot(aes(x=method, y=duration)) + 
    ggbeeswarm::geom_quasirandom() +
    stat_summary(fun.data = "mean_cl_boot", colour = "red", geom = "crossbar", size=0.4) +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) +
    scale_x_discrete("", labels=c(kmode="k-mode", mixdirdp100="mixdir DP\n(α₁=100)", mixdirdp="mixdir DP", mixdir="mixdir", lca="poLCA", rock="ROCK")) +
    scale_y_continuous(breaks=c(0.5, 1, 5, 10, 30, 60, 300, 600, 1800, 3600), minor_breaks = seq(1,10000, by=10),
                       labels=c("500 ms.", "1 s", "5 s", "10 s", "30 s", "1 min", "5 min", "10 min", "30 min", "1 h")) +
    facet_wrap(~ country) +
    ggtitle("Clustering Duration (K=25)") +
    coord_trans(y="log")
```


Create the combined figure

```{r fig.height=10, fig.width=12}
plot_grid(p_perf_lat5, p_perf_lat25, p_dur_lat5, p_dur_lat25, nrow=2, labels = c("A", "B", "C", "D"))

ggsave("../output/young_survey_cluster_consistency_all4.png", width=12, height=10)
```








# Compare performance on Ethiopia dataset (Fig. 7)


```{r fig.height=4.5, fig.width=7}
p_eth_5 <- class_assignment %>%
  filter(country == "Ethiopia") %>%
  filter(method != "mixdirdp100") %>%
  mutate(class = as.integer(as.factor(class))) %>%
  group_by(method, iter) %>%
  do({
    tmp <- .
    cross_df(list(a=unique(tmp$round), b=unique(tmp$round))) %>%
      filter(a < b) %>%
      rowwise() %>%
      mutate(ARI= mcclust::arandi(tmp[tmp$round == a, ]$class,
                                  tmp[tmp$round == b, ]$class))
  }) %>%
  ungroup() %>%
  mutate(method= fct_reorder(as.factor(method), -ARI, median)) %>%
  ggplot(aes(x=method, y=ARI)) + 
    ggbeeswarm::geom_quasirandom() +
    stat_summary(fun.data = "mean_cl_boot", colour = "red", geom = "crossbar", size=0.4) +
    ggsignif::geom_signif(comparisons = list(c("lca", "mixdir"), c("mixdir", "mixdirdp"), c("mixdirdp", "kmode")),
                        y_position=c(0.82, 0.74, 0.6), map_signif_level = TRUE, test.args=list(paired=FALSE),
                        textsize = 4) +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) +
    scale_x_discrete("", labels=c(kmode="k-mode", mixdirdp="mixdir DP",mixdirdp100="mixdir DP\n(α₁=100)", mixdir="mixdir", lca="poLCA", rock="ROCK")) +
    scale_y_continuous(limits=c(NA, 0.9), breaks=c(0, 0.25, 0.5, 0.75, 1)) +
    ggtitle("Clustering Consistency (K=5)") 

p_eth_25 <- class_assignment25 %>%
  filter(country == "Ethiopia") %>%
  filter(method != "mixdirdp100") %>%
  mutate(class = as.integer(as.factor(class))) %>%
  group_by(method, iter) %>%
  do({
    tmp <- .
    cross_df(list(a=unique(tmp$round), b=unique(tmp$round))) %>%
      filter(a < b) %>%
      rowwise() %>%
      mutate(ARI= mcclust::arandi(tmp[tmp$round == a, ]$class,
                                  tmp[tmp$round == b, ]$class))
  }) %>%
  ungroup() %>%
  mutate(method= fct_reorder(as.factor(method), -ARI, median)) %>%
  ggplot(aes(x=method, y=ARI)) + 
    ggbeeswarm::geom_quasirandom() +
    stat_summary(fun.data = "mean_cl_boot", colour = "red", geom = "crossbar", size=0.4) +
    ggsignif::geom_signif(comparisons = list(c("lca", "kmode"), c("lca", "mixdir"), c("mixdir", "mixdirdp")),
                          y_position = c(0.4, 0.5, 0.6), map_signif_level = TRUE, test.args=list(paired=FALSE),
                          textsize = 4) +
    ggsignif::geom_signif(comparisons = list(c("mixdirdp100", "mixdir")),
                          y_position = 0.7, map_signif_level = TRUE, test.args=list(paired=FALSE),
                          textsize = 4) +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) +
    scale_x_discrete("", labels=c(kmode="k-mode", mixdirdp="mixdir DP",mixdirdp100="mixdir DP\n(α₁=100)", mixdir="mixdir", lca="poLCA", rock="ROCK")) +
    scale_y_continuous(limits=c(NA, 0.9), breaks=c(0, 0.25, 0.5, 0.75, 1)) +
    ggtitle("Clustering Consistency (K=25)") 

plot_grid(p_eth_5, p_eth_25, labels=c("A", "B"))
cowplot::ggsave("../output/cluster_consistency.png", width=7, height=4.5)
```




# Consistency Alluvial Plot (Fig. 6)



```{r}
res1234 <- read_rds("../data/young_lives_clustering_results/young_lives_res1234_resrep.RDS")
res1 <- res1234[[1]]
res2 <- res1234[[2]]
res3 <- res1234[[3]]
res4 <- res1234[[4]]
res1_rep <- res1234[[5]]
```



```{r}

{par(
  fig=c(0, 0.5, 0.3, 1), # Set plot in upper half
  oma=c(0, 4, 1, 2) # Increase margins left and right
)

map2_df(res1_rep, 1:5, ~ data.frame(class=.x$pred_class, repetition=.y, ind=1:length(.x$pred_class))) %>%
  mutate(repetition=as.factor(paste0("Rep", repetition)),
         class=as.character(class)) %>%
  spread(repetition, class) %>%
  group_by_at(vars(starts_with("Rep")))  %>%
  summarize(freq=n()) %>%
  ungroup() %>%
  mutate(Rep1= case_when(
    Rep1 == "4" ~ "1",
    Rep1 == "1" ~ "3",
    Rep1 == "3" ~ "5",
    Rep1 == "5" ~ "2",
    TRUE ~ Rep1
  )) %>%
  mutate(Rep2= case_when(
    Rep2 == "2" ~ "5",
    Rep2 == "3" ~ "3",
    Rep2 == "4" ~ "2",
    TRUE ~ Rep2
  )) %>%
  mutate(Rep3= case_when(
    Rep3 == "5" ~ "1",
    Rep3 == "1" ~ "3",
    Rep3 == "3" ~ "5",
    Rep3 == "2" ~ "4",
    Rep3 == "4" ~ "2",
    TRUE ~ Rep3
  )) %>%
  mutate(Rep4= case_when(
    Rep4 == "1" ~ "1",
    Rep4 == "4" ~ "5",
    Rep4 == "5" ~ "4",
    TRUE ~ Rep4
  )) %>%
  mutate(Rep5= case_when(
    Rep5 == "3" ~ "1",
    Rep5 == "1" ~ "3",
    Rep5 == "2" ~ "5",
    Rep5 == "5" ~ "2",
    TRUE ~ Rep5
  )) %>%
  mutate(Rep1=LETTERS[6-as.numeric(Rep1)]) %>%
  mutate(Rep2=LETTERS[6-as.numeric(Rep2)]) %>%
  mutate(Rep3=LETTERS[6-as.numeric(Rep3)]) %>%
  mutate(Rep4=LETTERS[6-as.numeric(Rep4)]) %>%
  mutate(Rep5=LETTERS[6-as.numeric(Rep5)]) %>%
  {alluvial::alluvial(.[, 1:5], freq=.$freq, hide=.$freq < 0, blocks=TRUE,
                      gap.width = 0.2, alpha=0.8, axis_labels = rep("", 5))}


par(
  fig=c(0, 0.5, 0, 0.3), # Set plot to lower half of plot
  oma=c(1, 0.4, 0, 0), # Decrease outer margins left and right
  mar=c(3, 4.1, 1, 0), # Decrease top inner margin
  new=TRUE # Add to existing plot (weird I know...)
)
plot(map_dbl(res1_rep, ~ tail(.x$convergence, n=1)), type="l", ylab="ELBO", xlab="", xaxt="n")
points(map_dbl(res1_rep, ~ tail(.x$convergence, n=1)), pch=16)
axis(1, at=1:5, labels=paste0("Rep", 1:5))


par(
  fig=c(0.53, 1, 0.3, 1), # Set plot to lower half of plot
  oma=c(0, 4, 1, 1), # Increase margins left and right
  # mar=c(3, 4.1, 1, 1.5), # Decrease top inner margin
  new=TRUE # Add to existing plot (weird I know...)
)
rbind(data.frame(class=res1$pred_class,
                 childid=seq_along(res1$pred_class),
               round="1"),
      data.frame(class=res2$pred_class,
                 childid=seq_along(res2$pred_class),
               round="2"),
      data.frame(class=res3$pred_class,
                 childid=seq_along(res3$pred_class),
               round="3"),
      data.frame(class=res4$pred_class,
                 childid=seq_along(res4$pred_class),
               round="4")) %>%
  mutate(class=as.character(class),
         round=as.factor(paste0("R", round))) %>%
  spread(round, class) %>%
  group_by_at(vars(starts_with("R"))) %>%
  summarize(freq=n()) %>%
  ungroup() %>%
  mutate(R1= case_when(
    R1 == "3" ~ "1",
    R1 == "1" ~ "3",
    TRUE ~ R1
  )) %>%
  mutate(R2=case_when(
    R2 == "4" ~ "3",
    R2 == "3" ~ "4",
    TRUE ~ R2
  )) %>%
  mutate(R3=case_when(
    R3 == "5" ~ "1",
    R3 == "4" ~ "2",
    R3 == "2" ~ "3",
    R3 == "3" ~ "4",
    TRUE ~ R3
  )) %>%
  mutate(R4=case_when(
    R4 == "5" ~ "3",
    R4 == "2" ~ "5",
    R4 == "3" ~ "2",
    TRUE ~ R4
  )) %>%
  mutate(R1=LETTERS[as.numeric(R1)]) %>%
  mutate(R2=LETTERS[as.numeric(R2)]) %>%
  mutate(R3=LETTERS[as.numeric(R3)]) %>%
  mutate(R4=LETTERS[as.numeric(R4)]) %>%
  {alluvial::alluvial(.[, 1:4], freq=.$freq, hide=.$freq < 0, blocks=TRUE,
                      gap.width = 0.2, alpha=0.8, axis_labels = c(2002, 2006, 2009, 2013))}
}
dev.copy(png, "../output/young_survey_cluster_consistency_raw.png", width=4500, height=2100, res=300)
dev.off()
```









# Create combined performance plot

This part assumes that you ran the synthetic_data.RMD notebook before, so that the plots `p1` and `p2` are 
available to create a merged performance plot

```{r fig.height=4.5, fig.width=12}
title_row <- plot_grid(ggdraw() + draw_text("Synthetic Data Performance", size=18), 
                       ggdraw() + draw_text("Young Lives Consistency", size=18),
                       nrow=1)
plot_row <- plot_grid(p1 + ggtitle("K=4"), p2 + ggtitle("K=8"), 
                      p_eth_5 + ggtitle("K=5"), p_eth_25 + ggtitle("K=25"),
                      labels=c("A", "B", "C", "D"), nrow=1, align="h")
plot_grid(title_row, plot_row, nrow=2, rel_heights=c(1,7))
cowplot::ggsave("../output/cluster_performance_merged.png", width=12, height=4.5)
```




























